\textbf{Too Beautiful to be Fake: Attractive Faces are Less Likely to be Judged as Artificially Generated}

### 2022-10-03 submission/manuscript.Rmd
minor updates

In the second part of this study, 109 pictures of neutral-expression faces from the validated American Multiracial Face Database (AMFD, [@chen2021broadening]) were presented to the participants for 500ms each, in a randomized order, following a fixation cross display (750 ms). After each stimulus presentation, ratings of *Trustworthiness* ("I find this person trustworthy") and *Familiarity* ("This person reminds me of someone I know") were collected using visual analog scales. Notably, as facial attractiveness is a multidimensional construct, encompassing evolutionary, sociocultural, biological as well as cognitive aspects [@rhodes2006evolutionary; @han2018beauty], we assessed attractiveness using 2 visual analog scales, measuring general *Attractiveness* ("I find this person attractive") and physical *Beauty* ("This face is good-looking").

In the last part of the study, participants were informed that about half of images previously seen were AI-generated (the instructions used a cover story which mentioned that the research was aimed at validating a new face generation algorithm). The same set of stimuli was displayed again for 500 ms in a new randomized order. This time, after each display, participants were asked to express their belief regarding the nature of the stimulus using a visual analog scale (with *Fake* and *Real* as the two extremes). The study was implemented using *jsPsych* [@de2015jspsych], and the exact instructions are available in the experiment code.

@@

The analysis was carried out using *R 4.2* [@RCoreTeam2022], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@insightArticle; @bayestestRArticle; @performanceArticle; @parametersArticle; @correlationArticle]. As all the details, scripts and complimentary analyses are open-access, the manuscript will focus on significant results. In particular, given the statistical volatility of some of the findings, which itself could have been a testament to the shifting nature and inter-individuality of our sense of reality, the present study will concentrate on results that provide very strong evidence against the null hypothesis ($p <.01$).

@@

The *Paranoid Thinking* dimension of *Persecutory Ideation* was positively related to the probability of faces being judged as real ($\beta = 1.87$, $95\% CI = [0.99, 3.54]$, $z = 1.93$, $p = .054$). However, a negative relationship between confidence judgements and having paranoid beliefs was found for faces judged as real ($\beta_{real} = -0.21$, $95\% CI = [-0.29, -0.14]$, $z = -5.63$, $p < .001$). Social anxiety levels (obtained from aggregating the 2 social anxiety scales - *SIAS-6* and *SPS-6*), on the other hand, was significantly negatively related to confidence in real judgements only ($\beta_{real} = -0.22$, $95\% CI = [-0.30, -0.14]$, $z = -5.29$, $p < .001$). Similarly, *Intolerance to Uncertainty* was also negatively associated with confidence judgements of faces being real ($\beta_{real} = -0.30$, $95\% CI = [-0.45, -0.15]$, $z = -3.88$, $p < .001$). Specifically, although a negative association was observed between confidence in real judgements and the uncertainty intolerance dimension of *Inhibitory Anxiety* ($\beta_{real} = -1.00$, $95\% CI = [-2.01, 0.02]$, $z = -1.93$, $p = 0.054$), predominantly positive links were observed with its dimension of *Prospective Anxiety* ($\beta_{real} = 1.43$, $95\% CI = [0.10, 2.76]$, $z = 2.10$, $p = .036$).

@@

Questions pertaining to the attitude towards AI were reduced to 3 dimensions through factor analysis, labelled AI-Enthusiasm (loaded by items expressing interest and excitement in AI development and applications), AI-Realness (loaded by items expressing positive opinions on the ability of AI to create realistic material), and AI-Danger (loaded by items expressing concerns on the unethical misuse of AI technology). As such, although overall attitudes towards AI was found to be significantly and negatively associated with judgement confidence for real faces only ($\beta_{real} = -0.16$, $95\% CI = [-0.2, -0.13]$, $z = 8.90$, $p < .001$), suggesting that individuals with highly positive attitudes are less confident in rating faces as real, the extracted latent factor of AI-Enthusiasm instead displayed a significant positive relationship with the confidence in both real and fake judgments ($\beta_{real} = 0.21$, $95\% CI = [0.02, 0.40]$, $z = 2.20$, $p = .028$; $\beta_{fake} = 0.31$, $95\% CI = [0.12, 0.50]$, $z = -8.90$, $p < 0.001$).

@@

This study aimed at investigating the effect of facial ratings (attractiveness, beauty, trustworthiness and familiarity) on simulation monitoring, i.e., on the belief that the stimulus is artificially generated. The most striking result, in our opinion, is that despite all the stimuli being real faces from the same database, all participants, when given the information, believed (to high degrees of confidence) that a significant proportion of them were fake. This finding is a testimony to both the current expectations regarding CGI technology in the population, as well as to the volatility of our sense of reality. It underlines the strong impact of prior expectations and information on reality beliefs. In fact, stimuli-related and participant-related characteristics accounted for less than 20\% of the beliefs variance, suggesting that a large part of it is associated with other subjective processes.

@@

<!-- Attractiveness: males -->
Alternatively, we found a significant positive linear relationship in male participants for both attractiveness and beauty on simulation monitoring, that we could interpret under an evolutionary lens. Specifically, males purportedly place more emphasis on facial attractiveness as a sign of reproductive potential, as compared with females, who tend to value characteristics signaling resource acquisition capabilities [@qi2022gender; @fink2006facial; @buunk2002age]. However, while it is possible that the evolutionary weight associated with attractiveness skewed the perceived saliency towards attractive faces, rendering them significantly more salient than unattractive faces and in turn distorted the relationship with simulation monitoring, future studies should test this saliency-based hypothesis by measuring constructs closer to salience and its effects, for instance using neuroimaging [@lou2015perceptual; @indovina2007dissociation] or physiological markers [in particular, heart rate deceleration, @skora2022functional].

@@

<!-- trustworthiness -->
Our results found a positive linear trend between trustworthiness and simulation monitoring for females only; suggesting trustworthiness ratings linearly increased their beliefs in realness. Given the evidence positing that faces presented as real were rated more trustworthy [@liefooghe2022faces; @hoogers2021effect; @balas2017], we expected such a link to be more clearly present for both genders. One of the underlying mechanisms that possibly contributed to this dimorphism could have been the increased risk-taking aversion reported in females [explained evolutionarily as a compromise to their reproductive potential, @van2020sex], to which perceived facial trustworthiness relates [@hou2019survival]. While this hypothesis justifies much of our findings, particularly considering female participants' confidence judgements for both fake and real faces were also highest when these faces were rated very trustworthy and untrustworthy, for the hypothesis to hold true, faces deemed untrustworthy should have appeared as even more salient (representing an evolutionary threat), and hence be judged as more real. As such, further studies are needed to investigate the causes of the increased simulation monitoring sensitivity to trustworthiness in females.

<!-- familiarity -->
Contrary to our hypothesis, familiarity was not found to be significantly related to simulation monitoring decisions. Interestingly, however, there was a significant relationship between familiarity and confidence judgements for males only, who rated highly familiar faces as real and unfamiliar faces as fake with greater confidence. Previous studies have typically shown females to be superior at face recognition [@mishra2019gender; @lewin2002sex], displaying higher levels of facial memory [@sommer2013sex] and exhibiting greater metacognitive awareness [@kaplan2012effect]. As such, it is possible that the confidence judgements of females are less malleable to variability in face familiarity. Indeed, metacognitive awareness has been associated with greater cognitive inflexibility [@mishra2019gender]. However, this has not been corroborated by recent research, which found gender differences in the opposite direction instead (males were significantly overconfident regardless of their performance accuracy and the experimental conditions) [@bailey2021gender; @herbst2020gender]. Furthermore, it is to be noted that the present study's distribution of familiarity ratings was strongly skewed, and only a low number of pictures was rated as highly familiar. As such, future studies should clarify this point by experimentally manipulating familiarity, for instance by modulating the amount of exposure to items before querying the simulation monitoring judgments.

<!-- inter-indiv correlates -->
Regarding the role of inter-individual characteristics in simulation monitoring tendencies, we found that participants with higher scores of honesty-humility, a trait related to an increased risk perception and aversion [@levidi2022understanding; @weller2012honest], displayed a lower confidence in their simulation monitoring judgments.

Notably, greater narcissistic tendencies in dimensions such as acclaim seeking and manipulativeness was found to be associated with a higher probability of judging the face stimuli as real. Indeed, this finding is in line with recent research suggesting that narcissists were more susceptible to fake news [@piksa2022cognitive; @sindermann2020short], possibly due to the fact that they were less likely to engage in cognitive strategies such as reflective thinking [@littrell2020overconfidently; @ahadzadeh2021social] and are more vigilant of their environment [@carolan2017searching; @grapsas2020and; @eddy2021self]. This hence implies that individuals with greater narcissistic tendencies could have perceived the face images as more salient and engaged in less thoughtful information-processing, leading to their increased perceptions of the displayed stimulus being real.

@@

<!-- Social Anxiety, Paranoia and Uncertainty Intolerance -->
Unsurprisingly, individuals with more persecutory ideations (a form of paranoia where one believes others intend to cause them harm, @buck2019capturing) were more likely to rate the faces presented as real relative to fake. Consistent with previous findings that persecutory delusions are predicted by a greater sense of presence in a VR environment [@freeman2005psychology], it is likely that people with higher levels of paranoia are more sensitive to social threat and thus perceive the face stimuli presented with greater salience and emotionality, hence increasing their beliefs of its realness [@fornells2015people; @freeman2003can]. However, higher levels of persecutory ideation, social anxiety and intolerance of uncertainty were related to lower confidence in real judgements. While such results appear contradictory, it could be explained through the lens of fictional reappraisal, an emotion regulation strategy in which the belief that an emotional stimulus is not real is used as a mean to down-regulate the associated emotional reactions [@sperduti2017distinctive; @sperduti2016paradox; @singer2021cognitive]. Indeed, previous studies report that individuals with paranoia and anxiety engaged in more affective processing of external ambiguous stimuli [@fornells2015people]. This result is also in line with prior findings suggesting that the likelihood of reality beliefs decreased when pictures were emotional, as compared to when they were neutral [@makowski2016phenomenal]. As such, given the positive correlation between paranoia, social anxiety and uncertainty intolerance [@morrison2014moderating; @counsell2017intolerance; @barnby2020paranoia], it is possible that individuals high in such traits have greater interpersonal sensitivity and perceived the ambiguous face images with greater emotionality, therefore having to engage in emotional regulation strategies that resulted in a lower confidence for real judgements.

<!-- In fact, the only trait that modulated the tendency to rate an image as fake or real was the social phobia score, whereby low values were associated with an increased tendency to judge faces as real. Social phobia being characterized as a heightened fear in social interaction contexts [@kolassa2006psychophysiological; @ran2017impact], the nature of the stimuli - pictures of faces - could play a role in this association. However, given the fact that neutral faces are perceived as more threatening by participants with high social phobia [@yoon2008interpreting; @staugaard2010threatening; @peschard2017overestimation], and thus likely more salient, we would have expected high social phobia tendencies to be associated with a bias towards reality judgment. This apparent contradiction could be explained through the lens of fictional reappraisal,  **TODO: can you double check that this last statement is correct?**. Future studies should clarify the role of emotions and attempt at delineating processes related to emotion perception and reaction from emotion regulation. -->
<!-- However, it is important to note that all of the aforementioned dispositional traits did only influence the confidence judgments, but not the simulation monitoring bias *per se*. -->

<!-- Re-exposure Delay -->
On a methodological level, although the order of presentation of the facial images was randomized to reduce effects of adaptation, participants were more confident in their judgements for faces perceived as real following a shorter re-exposure delay. This may be because shorter re-exposure delays led to the faces being better remembered and appearing more familiar than faces that were displayed after a longer delay, thereby triggering self-referential and autobiographical memory processing during the repeated display [@abraham2009reality; @gobbini2013prioritized; @taylor2009neural]. In fact, this finding is in line with studies that show an up-regulation of emotions when fictional stimuli were associated with familiarity [@sperduti2016paradox; @makowski2017being], thus biasing its salience and perceived realness. However, that being the case, we would have expected shorter re-exposure delays to be related to a higher probability of the faces being rated as real. As such, future studies should further investigate the modulatory effects of types and degrees of familiarity on perceived realness judgements.

@@

In conclusion, research on our ability to discriminate between what is real from what is not is increasingly relevant given the current expansion on technologies related to virtual and augmented realities. The aim of the present study was to examine whether facial attractiveness can significantly influence our simulation monitoring decisions. In line with our hypothesis, we found faces rated as attractive and unattractive to be perceived as more real for females as a result of an increased salience. However, for males, the observation of a positive linear relationship between attractiveness and beauty with simulation monitoring decisions, adjusted for evolutionarily advantageous features such as trustworthiness, suggest a more complex relationship than hypothesized. As such, future studies could further investigate the role of attractiveness in simulation monitoring, taking greater consideration of the former's multidimensionality.

>>>>

In the second part of this study, 109 images of neutral-expression faces from the validated American Multiracial Face Database (AMFD, [@chen2021broadening]) were presented to the participants for 500ms each, in a randomized order, following a fixation cross display (750 ms). After each stimulus presentation, ratings of *Trustworthiness* ("I find this person trustworthy") and *Familiarity* ("This person reminds me of someone I know") were collected using visual analog scales. Notably, as facial attractiveness is a multidimensional construct, encompassing evolutionary, sociocultural, biological as well as cognitive aspects [@rhodes2006evolutionary; @han2018beauty], we assessed attractiveness using 2 visual analog scales, measuring general *Attractiveness* ("I find this person attractive") and physical *Beauty* ("This face is good-looking").

In the last part of the study, participants were informed that about half of the images previously seen were AI-generated (the instructions used a cover story which mentioned that the research was aimed at validating a new face generation algorithm). The same set of stimuli was displayed again for 500 ms in a new randomized order. This time, after each display, participants were asked to express their belief regarding the nature of the stimulus using a visual analog scale (with *Fake* and *Real* as the two extremes). The study was implemented using *jsPsych* [@de2015jspsych], and the exact instructions are available in the experiment code.

@@

The analysis was carried out using *R 4.2* [@RCoreTeam2022], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@insightArticle; @bayestestRArticle; @performanceArticle; @parametersArticle; @correlationArticle]. As all the details, scripts and complimentary analyses are open-access, and given that some of the results were statistically volatile, possibly due to the shifting nature and subjectivity of our sense of reality, this manuscript will concentrate on findings that are highly statistically significant ($p <.01$).

@@

A negative correlation was found between *Paranoid Thinking* and confidence judgements for faces judged as real ($\beta_{real} = -0.21$, $95\% CI = [-0.29, -0.14]$, $z = -5.63$, $p < .001$). Additionally, its dimension of *Persecutory Ideation* was positively related to the belief of the faces being real ($\beta = 1.87$, $95\% CI = [0.99, 3.54]$, $z = 1.93$, $p = .054$). Alternatively, overall social anxiety levels (obtained from aggregating the 2 social anxiety scales - *SIAS-6* and *SPS-6*) was significantly negatively related to confidence judgements, but only for faces judged as real ($\beta_{real} = -0.22$, $95\% CI = [-0.30, -0.14]$, $z = -5.29$, $p < .001$). Similarly, *Intolerance to Uncertainty* was also negatively associated with confidence judgements of faces being real ($\beta_{real} = -0.30$, $95\% CI = [-0.45, -0.15]$, $z = -3.88$, $p < .001$). In contrast, although a negative association was observed between confidence in real judgements and the uncertainty intolerance dimension of *Inhibitory Anxiety* ($\beta_{real} = -1.00$, $95\% CI = [-2.01, 0.02]$, $z = -1.93$, $p = 0.054$), predominantly positive links were observed with the dimension of *Prospective Anxiety* ($\beta_{real} = 1.43$, $95\% CI = [0.10, 2.76]$, $z = 2.10$, $p = .036$).

@@

Questions pertaining to the attitude towards AI were reduced to 3 dimensions through factor analysis, labelled AI-Enthusiasm (loaded by items expressing interest and excitement in AI development and applications), AI-Realness (loaded by items expressing positive opinions on the ability of AI to create realistic material), and AI-Danger (loaded by items expressing concerns on the unethical misuse of AI technology). As such, although overall attitudes towards AI was significantly and negatively associated with the judgement confidence for real faces ($\beta_{real} = -0.16$, $95\% CI = [-0.2, -0.13]$, $z = 8.90$, $p < .001$), suggesting that individuals with highly positive attitudes are less confident in rating faces as real, the extracted factor of AI-Enthusiasm instead displayed a significant positive relationship with the confidence in both real and fake judgments ($\beta_{real} = 0.21$, $95\% CI = [0.02, 0.40]$, $z = 2.20$, $p = .028$; $\beta_{fake} = 0.31$, $95\% CI = [0.12, 0.50]$, $z = -8.90$, $p < 0.001$).

@@

This study aimed at investigating the effect of facial ratings (attractiveness, beauty, trustworthiness and familiarity) on simulation monitoring, i.e., on the belief that the stimulus is artificially generated. The most striking result, in our opinion, is that despite all the stimuli being real faces from the same database, all participants, when given the information, believed (to high degrees of confidence) that a significant proportion of them were fake. This finding could thus be seen as a testimony to both the current expectations regarding CGI technology in the population, as well as to the volatility of our sense of reality. It underlines the strong impact of prior expectations and information on reality beliefs. In fact, stimuli-related and participant-related characteristics accounted for less than 20\% of the beliefs variance, suggesting that a large part of it is associated with other subjective processes.

@@

<!-- Attractiveness: males -->
Alternatively, we found a significant positive linear relationship in male participants for both attractiveness and beauty on simulation monitoring that we could interpret under an evolutionary lens. Specifically, males purportedly place more emphasis on facial attractiveness as a sign of reproductive potential, as compared with females, who tend to value characteristics signaling resource acquisition capabilities [@qi2022gender; @fink2006facial; @buunk2002age]. However, while it is possible that the evolutionary weight associated with attractiveness skewed the perceived saliency towards attractive faces, rendering them significantly more salient than unattractive faces and in turn distorted the relationship with simulation monitoring, future studies should test this saliency-based hypothesis by measuring constructs closer to salience and its effects, for instance using neuroimaging [@lou2015perceptual; @indovina2007dissociation] or physiological markers [in particular, heart rate deceleration, @skora2022functional].

@@

<!-- trustworthiness -->
Our results found a positive linear trend between trustworthiness and simulation monitoring for females only; suggesting trustworthiness ratings linearly increased their beliefs in realness. Given the evidence positing that faces presented as real were rated more trustworthy [@liefooghe2022faces; @hoogers2021effect; @balas2017], we expected such a link to be more clearly present for both genders. One of the underlying mechanisms that possibly contributed to this dimorphism could have been the increased risk-taking aversion reported in females [explained evolutionarily as a compromise to their reproductive potential, @van2020sex], to which perceived facial trustworthiness relates [@hou2019survival]. While this hypothesis justifies much of our findings, particularly considering female participants' confidence judgements for both fake and real faces were also higher when these faces were rated very trustworthy and untrustworthy, for the hypothesis to hold true, faces deemed untrustworthy should have appeared as even more salient (representing an evolutionary threat), and hence be judged as more real. As such, further studies are needed to investigate the causes of the increased simulation monitoring sensitivity to trustworthiness in females.

<!-- familiarity -->
Contrary to our hypothesis, familiarity was not found to be significantly related to simulation monitoring decisions. Interestingly, there was a significant relationship between familiarity and confidence judgements for males only, who rated highly familiar faces as real and unfamiliar faces as fake with greater confidence. However, taking into perspective that previous studies report females to be superior at face recognition [@mishra2019gender; @lewin2002sex], displaying higher levels of facial memory [@sommer2013sex] and exhibiting greater metacognitive awareness [@kaplan2012effect], it is possible that the confidence judgements of females are less malleable to variability in face familiarity. Indeed, metacognitive awareness has been associated with overconfidence in making judgements [@mata2013metacognitive; @klein2018metacognitive]. However, it should be noted that this has not been corroborated by recent research, which found males were significantly overconfident regardless of their performance accuracy and the experimental conditions [@bailey2021gender; @herbst2020gender]. Furthermore, the present study's distribution of familiarity ratings was strongly skewed, and only a low number of pictures was rated as highly familiar. As such, future studies should clarify this point by experimentally manipulating familiarity, for instance by modulating the amount of exposure to items before querying the simulation monitoring judgments.

<!-- inter-indiv correlates -->
Regarding the role of inter-individual characteristics in simulation monitoring tendencies, we found that participants with higher scores of honesty-humility, a trait related to an increased risk perception and aversion [@levidi2022understanding; @weller2012honest], displayed a lower confidence in their simulation monitoring judgments. Notably, greater narcissistic tendencies in dimensions such as acclaim seeking and manipulativeness was found to be associated with a higher probability of judging the face stimuli as real. Indeed, this finding is in line with recent research suggesting that narcissists were more susceptible to fake news [@piksa2022cognitive; @sindermann2020short], possibly due to the fact that they were less likely to engage in cognitive strategies such as reflective thinking [@littrell2020overconfidently; @ahadzadeh2021social] and are more vigilant to external stimuli [@carolan2017searching; @grapsas2020and; @eddy2021self]. This hence implies that individuals with greater narcissistic tendencies could have perceived the face images as more salient and engaged in less analytic-processing, leading to their increased perceptions of the displayed stimulus being real.

@@

<!-- Social Anxiety, Paranoia and Uncertainty Intolerance -->
Unsurprisingly, individuals with more persecutory ideations (a form of paranoia where one believes others intend to cause them harm, @buck2019capturing) were more likely to rate the faces presented as real relative to fake. Consistent with previous findings that persecutory delusions are predicted by a greater sense of presence in a VR environment [@freeman2005psychology], it is likely that people with higher levels of paranoia are more sensitive to social threat and thus perceive the face stimuli presented with greater salience and emotionality, hence increasing their beliefs of its realness [@fornells2015people; @freeman2003can]. However, higher levels of persecutory ideation, social anxiety and intolerance of uncertainty were related to lower confidence in real judgements. While these results appear contradictory, such findings could be explained through the lens of fictional reappraisal, an emotion regulation strategy in which the belief that an emotional stimulus is not real is used as a means to down-regulate the associated emotional reactions [@sperduti2017distinctive; @sperduti2016paradox; @singer2021cognitive]. Previous studies report that individuals with paranoia and anxiety engaged in more affective processing of external ambiguous stimuli [@fornells2015people]. This result is also in line with prior findings which suggest that the likelihood of reality beliefs decreased when pictures were emotional, as compared to when they were neutral [@makowski2016phenomenal]. As such, it is possible that individuals high in such anxiety-related traits are more sensitive to social cues and perceived the ambiguous face images with greater emotionality, therefore having to engage in emotional regulation strategies that resulted in a lower confidence for real judgements.

<!-- given the known positive correlations between paranoia, social anxiety and uncertainty intolerance [@morrison2014moderating; @counsell2017intolerance; @barnby2020paranoia] -->
<!-- In fact, the only trait that modulated the tendency to rate an image as fake or real was the social phobia score, whereby low values were associated with an increased tendency to judge faces as real. Social phobia being characterized as a heightened fear in social interaction contexts [@kolassa2006psychophysiological; @ran2017impact], the nature of the stimuli - pictures of faces - could play a role in this association. However, given the fact that neutral faces are perceived as more threatening by participants with high social phobia [@yoon2008interpreting; @staugaard2010threatening; @peschard2017overestimation], and thus likely more salient, we would have expected high social phobia tendencies to be associated with a bias towards reality judgment. This apparent contradiction could be explained through the lens of fictional reappraisal,  **TODO: can you double check that this last statement is correct?**. Future studies should clarify the role of emotions and attempt at delineating processes related to emotion perception and reaction from emotion regulation. -->
<!-- However, it is important to note that all of the aforementioned dispositional traits did only influence the confidence judgments, but not the simulation monitoring bias *per se*. -->

<!-- Re-exposure Delay -->
On a methodological level, although the order of presentation of the facial images was randomized to reduce effects of adaptation, participants were more confident in their judgements for faces perceived as real following a shorter re-exposure delay. This may be because shorter re-exposure delays led to the faces being better remembered and appearing more familiar than faces that were displayed after a longer delay, thereby triggering self-referential and autobiographical memory processing during the repeated display [@abraham2009reality; @gobbini2013prioritized; @taylor2009neural]. In fact, this finding is in line with studies that show an up-regulation of emotions when fictional stimuli were associated with familiarity [@sperduti2016paradox; @makowski2017being], thus biasing its salience and perceived realness. However, that being the case, we would have expected shorter re-exposure delays to be related to a greater belief of the faces being real. As such, future studies should further investigate the modulatory effects of types and degrees of familiarity on perceived realness judgements.

@@

In conclusion, research on our ability to discriminate between what is real from what is not is increasingly relevant given the current expansion on technologies related to virtual and augmented realities. The aim of the present study was to examine whether facial attractiveness significantly influences our simulation monitoring decisions. In line with our hypothesis, we found faces rated as attractive and unattractive to be perceived as more real for females as a result of an increased salience. However, for males, the observation of a positive linear relationship between attractiveness and beauty with simulation monitoring decisions, adjusted for evolutionarily advantageous features such as trustworthiness, suggest a more complex relationship than hypothesized. As such, future studies could take greater consideration of the multidimensionality of facial attractiveness in investigating its role in simulation monitoring decisions.
